The CrypTool Script: \\ Cryptography, Mathematics and more \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free eLearning program CrypTool\\ (8th edition -- distributed with CrypTool version 1.4.10)} \vspace{60pt}

### 2007-05-31 script/en/chapters/primes.tex
Script: Überblick + Vorwort aktualisiert

\hline
\end{tabular}
\caption{The 20 largest known primes and its particular number types
         (as of May 2007)}    % Eyecatcher_neue-Mersenne
\end{center}
\end{table} 
\footnotetext{\index{Number!Sierpinski}\index{Seventeen or Bust SoB}%
This number was found within the distributed computing project
``Seventeen or Bust'' (SoB) (\href{http://www.seventeenorbust.com}{\texttt{http://www.seventeenorbust.com}})
% {\href{http://www.mersenne.org} {\tt http://www.mersenne.org}}
at March 26, 2007. While the well known \hyperlink{GIMPS-project}{GIMPS project} (chapter~\ref{zahlentyp_mersenne}) searches for bigger and bigger of the infinitely many primes, there is a chance, that the SoB project could have been completed its
task sometime.

@@

An easy way of calculating all $PI(x)$ primes less than or equal to $x$ is to
use the sieve of Eratosthenes. In the 3rd century B.C., he found an
extremely easy, automatic way of finding this out. To begin with, you write
down all numbers from 2 to $x$, circle 2, then cross out all multiples of 2.
Next, you circle the lowest number that hasn't been circled or crossed out
(3) and again cross out all multiples of this number, etc. You only need to
continue until you reach the largest number whose square is less than or
equal to $x$.

@@

\vskip +20pt 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|}\hline
Probability that you will be hijacked on your next flight	& $ 5.5 \cdot 10^{-6} $ \\
Annual probability of being hit by lightning 	& $ 10^{-7} $ \\
Probability of 6 correct numbers in the lottery & $ 7.1 \cdot 10^{-8} $ \\
Risk of being hit by a meteorite & $ 1.6 \cdot 10^{-12} $  \\
\hline
Time until the next ice age (in years) & $14,000 $ = $ (2^{14})$ \\
Time until the sun dies (in years)   & $10^{9} $ =  $(2^{30})$ \\
Age of the earth (in years) & $ 10^9 $  =  $(2^{30}) $  \\
Age of the universe (in years) & $ 10^{10} $ =  $(2^{34}) $ \\
Number of the earth's atoms    & $10^{51} $ =  $ (2^{170}) $ \\
Number of the sun's atoms      & $10^{57}$  =  $ (2^{190})$ \\
Number of atoms in the universe (without dark material)  & $10^{77}$ =  $ (2^{265})$ \\
Volume of the universe (in $cm^3$)   & $10^{84}$ = $(2^{280})$ \\ \hline
\end{tabular}
\caption{Likelihoods from physics and everyday life (dimensions)}
\end{center}
\end{table}

>>>>

\hline
\end{tabular}
\caption{The 20 largest known primes and its particular number types
         (as of May 2007)}    % Eyecatcher_neue-Mersenne
\label{L_n_Largest_Kown-Primes}
\end{center}
\end{table} 
\footnotetext{\index{Number!Sierpinski}\index{Seventeen or Bust SoB}%
This number was found within the distributed computing project
``Seventeen or Bust'' (SoB) (\href{http://www.seventeenorbust.com}{\texttt{http://www.seventeenorbust.com}})
% {\href{http://www.mersenne.org} {\tt http://www.mersenne.org}}
at March 26, 2007. While the well known \hyperlink{GIMPS-project}{GIMPS project} (chapter~\ref{zahlentyp_mersenne}) searches for bigger and bigger of the infinitely many primes, there is a chance, that the SoB project could have been completed its
task sometime.

@@

An easy way of calculating all $PI(x)$ primes less than or equal to $x$ is to
use the sieve of Eratosthenes. In the 3rd century B.C., he found an
extremely easy, automatic way of finding this out. To begin with, you write
down all numbers from 2 to $x$, circle 2, then cross out all multiples of 2.
Next, you circle the lowest number that hasn't been circled or crossed out
(3) and again cross out all multiples of this number, etc. You only need to
continue until you reach the largest number whose square is less than or
equal to $x$.\index{NT, Learning Tool for Number Theory}%
\footnote{%
    With the educational tool for number theory {\bf NT} you can apply the
    sieve of Eratosthenes in a computer-aided and guided way: Enter you own
    number and do the sieving step by step: See learning unit 1.2,
    pages 6/21 and 7/21.\\
    NT can be called in CrypTool\index{CrypTool} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
     \textbackslash{} Learning Tool for Number Theory}.
}

@@

\vskip +20pt 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|}\hline
Probability that you will be hijacked on your next flight	& $ 5.5 \cdot 10^{-6} $ \\
Annual probability of being hit by lightning 	& $ 10^{-7} $ \\
Probability of 6 correct numbers in the lottery & $ 7.1 \cdot 10^{-8} $ \\
Risk of being hit by a meteorite & $ 1.6 \cdot 10^{-12} $  \\
\hline
Time until the next ice age (in years) & $14,000 $ = $ (2^{14})$ \\
Time until the sun dies (in years)   & $10^{9} $ =  $(2^{30})$ \\
Age of the earth (in years) & $ 10^9 $  =  $(2^{30}) $  \\
Age of the universe (in years) & $ 10^{10} $ =  $(2^{34}) $ \\
Number of molecules within one waterdrop &  $10^{20} $  =  $(2^{63})$ \\
Number of the earth's atoms    & $10^{51} $ =  $ (2^{170}) $ \\
Number of the sun's atoms      & $10^{57}$  =  $ (2^{190})$ \\
Number of atoms in the universe (without dark material)  & $10^{77}$ =  $ (2^{265})$ \\
Volume of the universe (in $cm^3$)   & $10^{84}$ = $(2^{280})$ \\ \hline
\end{tabular}
\caption{Likelihoods and dimensions from physics and everyday life}
\end{center}
\end{table}
